### 2020-02-06 Untitled.Rmd
Switched to non interaction model
note that I only ran CS3 so the pdf is not complete. fixes include all red text

Jesse Still needs to clean up the text and deliberate his write up in the case study. Code will also be cleaned to be more readable.

---
title: "ipareport"
author: "Jesse"
date: "30/01/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r run}
library(casebase)
library(survival)
library(riskRegression)
data(veteran)

new_data <- data.frame(trt = "test",
                       celltype = "adeno",
                       karno = median(veteran$karno),
                       diagtime = median(veteran$diagtime),
                       age = median(veteran$age),
                       prior = "no")

modelKM <- coxph(Surv(time, status) ~ 1,
                  data = veteran, y=TRUE, x = TRUE)
abKM<- survfit(modelKM)

modelCox <- coxph(Surv(time, status) ~ karno + diagtime + age + prior + celltype + trt,
                data = veteran, y=TRUE, x = TRUE)
abCox<- survfit(modelCox)

modelCB <- fitSmoothHazard(status ~ time + karno + diagtime + age + prior +
                              celltype + trt, data = veteran, ratio = 100)
abCB <- absoluteRisk(object = modelCB, time = seq(0,11, sep=99)) ##takes long to run

plot(abCB[,1],rowMeans(abCB[,-c(1)]),type='l',col="black",lwd=3,main="Support- CaseBase vs. Cox+glmnet vs KM Absolute Risk Curves",xlab="Survival-Time",ylab="Cumulative Incidence",ylim=c(0,1),xlim=)
lines(abCox,col="red",fun="event",lwd=3,conf.int = FALSE)
lines(abKM ,col="Blue",fun="event",lwd=3,conf.int = FALSE)
legend("bottomright",
       legend = c( "Casebase (Lasso+linear)","semi-parametric (Cox)+glmnet","KM curve"),
       col = c("black","red","Blue"),
       lty = c(1, 1, 1),
       bg = "gray90")

s=Score(list("KM"=modelKM,"Cox"=modelCox),
        formula=Surv(time,status)~1,data=veteran,conf.int=FALSE,times=c(5,8))

BrierNoCensoring=sum((1-rowMeans(abCB[,-c(1)]))^2)/length(abCB[,1])

fit=modelCox
brierErrorCox <- function(fit, Time,data) {
    cfit <- survfit(Surv(time, status==0)~1, data) #KM curve
    sfit <- survfit(fit, newdata = data)  # model curve for fit2
    S <- summary(sfit, Time)$surv
    cProb <- summary(cfit, Time)$surv
    Score <- matrix(NA, nrow(S), ncol(S))
    for (i in 1:length(Time)) {
        CensBefore <- fit$y[,2]==0 & fit$y[,1] < Time[i]
        y <- fit$y[,1] > Time[i]
            Score[i,] <- (y - S[i,])^2

        Score[i, CensBefore] <- 0
    }
    Err <- sweep(Score, 1, cProb, '/')
    apply(Err, 1, mean)
}
fit=modelCB
data=veteran
brierErrorCB <- function(fit, Time,data) {
    cfit <- survfit(Surv(time, status==0)~1, data) #KM curve
    sfit <- absoluteRisk(object = modelCB, time=Time)  # model curve for fit2
    S <- 1-sfit[,-c(1)]
    cProb <- summary(cfit, Time)$surv
    Score <- matrix(NA, nrow(S), ncol(S))
    for (i in 1:length(Time)) {
        CensBefore <- data$status==0 & fit$originalData$time < Time[i]
        y <- fit$originalData$time > Time[i]
        Score[i,] <- (y - S[i,])^2
        Score[i, CensBefore] <- 0
    }

    Err <- sweep(Score, 1, cProb, '/')
    Err=rowMeans(Err)

}

Time <- seq(0, 300, len=200)
errCox <- cbind(brierErrorCox(modelCox, Time,veteran))
errCB<- cbind(brierErrorCB(modelCB,Time,veteran))
errors<-data.frame(coxPredictionError=errCox,cbPredictionError=errCB)
library(ggplot2)
library(dplyr)
library(tidyr)
errors=na.omit(errors)  %>%
    gather(key,value, coxPredictionError, cbPredictionError)
errors$time=Time

errors%>%
    ggplot(aes(x=time,y=value,col=key,group=key))+geom_line(lwd=1)+

    labs(x = "time",y="Brier prediction error")

```